### 2018-06-07 Writing/response_to_reviewers.tex

\item{\textcolor{blue}{Along the lines set out by Reviewer 2, it would be wise to ensure that your metric is characterised as thoroughly as possible. In this regard, does it meet the following five conditions?}}

\begin{enumerate}
\item{\textcolor{blue}{Distinctiveness}}
\item{\textcolor{blue}{Non-negativity}}
\item{\textcolor{blue}{Symmetry}}
\item{\textcolor{blue}{Triangle inequality}}
\item{\textcolor{blue}{Four-point condition}}
\end{enumerate}

\item{\textcolor{blue}{If so, then it is actually an evolutionary distance, and it can be used as such; if not, then it is another type of distance. Establishing the type of distance may be overkill, but the characterization of a metric helps to determine what it can be used for, and what it cannot be used for.}}

In the supplementary materials 1, we mathematically define the characteristics of $CD$ and proved that it is a distance metric since it respects the four first points:

\begin{enumerate}
\item Distinctiveness: $CD_{(x,y)} = 0 \Leftrightarrow x = y$, proved in equation 10 (note that we corrected the error highlighted by reviewer 2 with a more restrictive definition of $x$ and $y$).
\item Non-negativity: $CD_{(x,y)} \geq 0$, proved in equation 9.
\item Symmetry: $CD_{x,y} = CD_{y,x}$, proved in equation 11.
\item Triangle inequality: $CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}$, proved in equation 12.
\end{enumerate}

We did not found a proof for the Four-point condition that satisfied the equation:

$CD_{(i,j)} + CD_{(m,n)} \leq \text{max}{(CD_{(i,m)} + CD_{(j,n)} ; CD_{(j,m)} + CD_{(i,n)} )}$

So we cannot confirm or reject that $CD$ is a evolutionary distance metric.

@@

\item{\textcolor{blue}{I feel that the manuscript is certainly worth publishing in Systematic Biology, although I worry that the simulation procedure used by the authors limits their ability to thoroughly test of the effect of between character correlation on the accuracy of estimating phylogenetic trees. The authors note this themselves in the discussion, but I feel more should be made of this point. If the simulations utilised an explicit distribution of character covariance then I think that interpretation of the results would have been more straightforward.}}

%TG: Not sure how to deal with this one.

\item{\textcolor{blue}{Another worry I have is that the implied covariance at the tips does not necessarily reflect covariation over the full phylogeny. It is conceivable that two characters can appear to have near perfect correlation at the tips, but arrived at these states through completely independent pathways, particularly when simulating with a continuous time Markov chain where unobserved transitions to a new state and back to the original state are possible.}}

%TG: Not sure about this one either: I guess its impossible to say both in simulations or in real life: in the end we only have the tips and their information.

@@

\item{\textcolor{blue}{It seems that the results reported by the authors are inherently driven by the phylogenetic signal present in their ``bootstrapped'' datasets. This effect would be present in any approach to the simulation of covarying characters, but as the authors do not explicitly simulate covariance but instead rely on a proxy they cannot be certain that their results are driven by true character covariation or purely a restructuring of phylogenetic signal by resampling. In a simulation framework in which covariance is actively enforced and controlled it would be possible to assess other aspects of character covariance that seem important for empirical datasets, such as correlation within morphological compartments. I think without tests of these sort we cannot be completely sure of the effect of character covariance on topology estimates. Despite this, the authors present an overview of the effect of character covariance at large, and I certainly think that this is a worthwhile contribution.}}

%TG: Maybe say something along the lines: "yeah, great idea we'll do it later"?

@@

\item{\textcolor{blue}{A discussion of the ``effective'' number of characters might also be warranted. When maximising character covariance this likely requires the resampling of the same handful of characters multiple times, each of which provides the exact same phylogenetic signal. The redundancy of these characters seems likely to influence the accuracy of estimates. The choice of characters that are sampled by the bootstrap procedure will therefore constrain the accuracy of topology estimates. If the resampled matrix consists of a handful of characters with high consistency with the true tree, sampled many times over, then the tree estimate will be good (for at least part of the tree) and covariance will be high. Similarly, if the handful of regularly resampled characters are mainly inconsistent (with the true tree) then tree estimates will be poor but the implied covariance will also be high. I think this is evident from the reduced performance in larger matrices, where there will be a significant reduction in phylogenetic signal when attempting to maximise covariance.}}

%TG: needs to emphasize this point: basically, smaller matrices matrices are more sensible to what the reviewer suggests (i.e. the chance of pick up the right or the wrong characters are higher). In bigger matrices, this variance is reduced because the number of good/bad characters is diffused.

@@

\item{\textcolor{blue}{Overall, I think that the authors provide an overview of the effect of covariance on the estimation of phylogeny, but without an explicit simulation of patterns of covariation we do not yet have a definitive understanding of the effects of between character covariance.}}

We definitely agree with this statement and are working on a follow up project with more explicit patterns of covariation.
However, one major bottleneck in such analysis is our ability to actually simulate discrete morphological characters (regardless of their pattern of covariance) that are meaningful in reflecting the actual patterns observed in empirical discrete morphological characters.
Thankfully, the debate between Goloboff et al. and O'Reilly et al. will provide us with some more insightful and realistic methods to simulate discrete morphological characters.

Secondly, as we highlighted throughout the manuscript (lines @@@), the main mechanisms that generate phylogenetic signal and variance/covariance between characters are the intra-organismal dependence and the evolutionary dependence described in the introduction.
Both of which can unfortunately only be observed/analysed, \textit{a posteriori} based on a phylogenetic hypothesis.

@@

\subsection{Minor revisions:}

\begin{enumerate}

\item{\textcolor{blue}{2:2 – Phylogenetic algorithms don't necessarily require independence, they simply assume it.}}

We've rephrased this sentence to:
\textit{Phylogenetic analysis algorithms assume character independence}. lines @@@

### 2018-06-07 Writing/draft_highlighted.tex
Some equation proofs

\newpage
\begin{abstract}
Phylogenetic analysis algorithms require the assumption of character independence - a condition generally acknowledged to be violated by morphological data.
Correlation between characters can originate from intra-organismal features, shared phylogenetic history or forced by particular character-state coding schemes. 
Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with \hl{careful character coding}, phylogenetic software do not distinguish between any of them.

>>>>

\newpage
\begin{abstract}
\hl{Phylogenetic analysis algorithms assume character independence} - a condition generally acknowledged to be violated by morphological data.
Correlation between characters can originate from intra-organismal features, shared phylogenetic history or forced by particular character-state coding schemes. 
Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with \hl{careful character coding}, phylogenetic software do not distinguish between any of them.
